---
title: "Example Document"
subtitle: "This is an example"
author: "Martin Reinhardt"
output: pdf_document
---

# Hello R World 

## Introduction

This is a very short document, which will show you everything which is possible with R Markdown.

## General Markdown syntax

* Headline:
  ```
  # Headline level 1
  ## Headline level 2
  ### Headline level 3
  #### Headline level 4
  ```
* *italics* `*italics*`
* _italics_ `_italics_`
* **bold** `**bold**`
* Sub~script~ `Sub~script~`
* Super^script^ `Super^script^`
* `inline code` 
  ```
  `inline code`
  ```
* [hyperlink](https://duckduckgo.com) `[hyperlink](https://duckduckgo.com)`
* <img src="https://cran.r-project.org/Rlogo.svg" style="height:18pt"> `![image](https://cran.r-project.org/Rlogo.svg)`

```
- list item 1
- list item 2

* list item 3
* list item 4
```

- list item 1
- list item 2

* list item 3
* list item 4

```
1. Numbered list item 1
1. Numbered list item 2
```

1. Numbered list item 1
1. Numbered list item 2

Math uses LaTeX notation:

```
$$f(x) = x^2$$
```

$$f(x) = x^2 + c$$

use html/latex code for special formating, this makes the document less portable:

```
<p style="color:green;">hello</p>
```

<p style="color:green;">hello</p>


## R in Markdown

````
`r ''````{r}
f <- function(x) {
  sum(x ^ 2)
}
```
````

```{r}
f <- function(x) {
  sum(x ^ 2)
}
```

The session is saved in background where the process was startet:

```{r}
f(1:5)
```

This also works with plots:

````
`r ''````{r}
plot(1:10)
```
````

```{r}
plot(1:10)
```

### chunk options

show code without evaluating

````
`r ''````{r, eval = FALSE}
x <- 1:10
x
```
````

```{r, eval = FALSE}
x <- 1:10
x
```

show only the results

````
`r ''````{r, echo = FALSE}
x <- 1:10
x
```
````

```{r, echo = FALSE}
x <- 1:10
x
```